Casual Bandits

### 2016-01-11 causal-bandit/icml-2015/causal-bandits-icml.tex
minor fixes plus experiment details

The idea is best illustrated with an example. Consider a farmer wishing to optimise the yield of her crop. In each season she can invest in a green house (to control
the temperature), cover her plants with nets (to prevent loss to birds), use expensive fertilizers, or install a high-tech watering system. We will assume for
simplicity that she does not have the resources to make more than one intervention. Of course, the watering system will only be useful if the rainfall is low and
the nets will be useful if the bird population is large. These things are random events that change from season to season, which makes the impact of each intervention
difficulty to measure. The goal of this paper is to design an optimal strategy for the farmer to learn which intervention will be most profitable.

@@

@@

\todot{this remark feels out of place, but where to put it?}
\begin{remark}
In order to be consistent with the literature on causal inference we use the notation $do()$ to denote the action of doing nothing and $do(X_{t,I_t} = J_t)$
the action of intervening on the $I_t$th variable and setting it to equal $J_t$. 
In the bandit community it is implicit that 
algorithms selecting actions are intervening in the system. So it is sufficient to index actions according to the variable and value. 
However, in causal graphs, it is essential to differentiate observing (or conditioning) on a variable taking a certain value, from 
intervening to set that variable. Although in the specific causal graph we consider, observation and intervention are the same, we 
deliberately introduce the do-notation \cite{Pearl2000} that makes this distinction clear so as to help provide a bridge between the 
bandit and causal inference communities.
\end{remark}

@@

We propose a simple explore-exploit based algorithm that leverages  \eqref{eq:observe}. Without loss of generality, we assume $q_i \in [0,\frac{1}{2}]$ and $q_1 \leq q_2 ... \leq q_N$.

@@

Algorithm \ref{alg:known_q} relies on $\boldsymbol{q}$ and thus $m$ being known. We now consider the case where $\boldsymbol{q}$ is unknown. We begin by considering the simple regret, defined as the expected difference between the mean payoff of the optimal action and that of the action estimated to be optimal in $T$th round.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EXPERIMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}

>>>>

The challenge is best illustrated with an example. Consider a farmer wishing to optimise the yield of her crop. In each season she can invest in a green house (to control
the temperature), cover her plants with nets (to prevent loss to birds), use expensive fertilizers, or install a high-tech watering system. We will assume for
simplicity that she does not have the resources to make more than one intervention. Of course, the watering system will only be useful if the rainfall is low and
the nets will be useful if the bird population is large. These things are random events that change from season to season, which makes the impact of each intervention
difficulty to measure. The goal of this paper is to design an optimal strategy for the farmer to learn which intervention will be most profitable.

@@

\begin{remark}
In order to be consistent with the literature on causal inference we use the notation $do()$ to denote the action of doing nothing and $do(X_{t,I_t} = J_t)$
the action of intervening on the $I_t$th variable and setting it to equal $J_t$. 
In the bandit community it is implicit that 
algorithms selecting actions are intervening in the system. So it is sufficient to index actions according to the variable and value. 
However, in causal graphs, it is essential to differentiate observing (or conditioning) on a variable taking a certain value, from 
intervening to set that variable. Although in the specific causal graph we consider, observation and intervention are the same, we 
deliberately introduce the do-notation \cite{Pearl2000} that makes this distinction clear so as to help provide a bridge between the 
bandit and causal inference communities.
\end{remark}

@@

@@

We begin by considering the case where $\boldsymbol{q}$ is known and propose a simple explore-exploit based algorithm that leverages  \eqref{eq:observe}. Without loss of generality, we assume $q_i \in [0,\frac{1}{2}]$ and $q_1 \leq q_2 ... \leq q_N$.

@@

We now demonstrate we can obtain the same order of regret for the case where $\boldsymbol{q}$ is unknown.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EXPERIMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
In all experiments, the reward is generated according to:
\eq{
Y_t \sim \begin{cases}
\bernoulli(.5+\epsilon) & \text{if } X_1 = 1 \\
\bernoulli(\frac{.5-(.5+\epsilon)q_1}{1-q_1}) & \text{otherwise}\,.
\end{cases}
}

With $\epsilon \in (0,.5)$. This ensures that 
\eq {
\P{Y|do(X_i = j)} = \begin{cases}
.5+\epsilon & \text{if } (i,j)=(1,1) \\
\frac{.5-(.5+\epsilon)q_1}{1-q_1} & \text{if } (i,j) = (1,0) \\
.5 & \text{otherwise}
\end{cases}
}

This allows us to rapidly simulate rewards and conditional rewards for large numbers of variables and to change the value of $\boldsymbol{q}$ with minimal impact on the reward function \footnote{$\boldsymbol{q}$ only effects the reward of a single, non-optimal arm, which becomes insignificant with large numbers of variables.}.
